Projection variability: Is the family of sentences really a family?

### 2024-02-14 writing/SuB-paper/projection-operators-SuB.tex
tracking udpates from overleaf wed afternoon

%% other packages I added (LH)
% misc formatting
	\usepackage{booktabs}
    \usepackage[export]{adjustbox}
    \usepackage{rotating}

@@

    Some research, however, suggests that entailment-cancelling operators may affect projection differentially. For instance, \citet{karttunen_observations_1971} proposed distinguishing English factive predicates (e.g., \textit{regret}) from semi-factives (e.g., \textit{discover}). Based on \Next, he suggested that the content of the complement (CC) of true factives consistently projects across the operators in \Last, but that of semi-factives does not always project from under polar questions, modals, or conditionals.%

@@

    %    The factive/semi-factive distinction (iii) is further called into question by \citealt{beaver_have_2010}, who reports many naturally occurring examples of cognitive factive predicates, showing that their CC does not necessarily project (e.g., English \emph{be/become aware, know, discover, notice, find out, realize, accept the fact}; French \emph{trouver} \lq discover\rq).

    %Experimental findings from \citet{djarv_cognitive_2018}, \citet{tonhauser_how_2018}, and \citet{degen_are_2022} can be evaluated with respect to whether they support Karttunen's distinction (iii), with mixed results. These studies observed by-predicate variation in polar questions, where more projection is expected for factives than semi-factives.
	%
	%\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, did find higher ratings for English \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive). This could be interpreted as limited support for (iii). But here, it is also not obvious how exactly this task relates to projection.\footnote{Say why}

	%\citet{tonhauser_how_2018} and \citet{degen_are_2022} measured speaker commitments toward the CC of a broad range of English clause-embedding predicates more directly, by collecting judgments of how certain participants take a speaker to be of the contents under investigation.
	%
	%The observed differences between the various predicates do not align with theoretically assumed lexical classes. While \citet{degen_are_2022} demonstrate this for the distinction between purported factive and non-factive predicates, the results did not match the expectations from Karttunen's factive vs. semi-factive classification either.
	%
	%\citet{tonhauser_how_2018} found, e.g., that the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}. \citet{degen_are_2022} observed that the CC of factive \textit{reveal} was less projective than that of semi-factive \textit{discover}, which, in turn, was less projective than that of factive \textit{know}. These findings also suggest that examining projection ratings for various predicates individually, rather than pre-grouping them based on theoretical assumptions, can provide more nuanced insights into by-predicate projection variation.

	%Both \citet{smith_relationship_2014} \citet{sieker_projective_2022} found that projection is affected differently by entailment-cancelling operators. However, while Smith \& Hall found variation in the effect of the operator on projection between various projective contents, Sieker \& Solstad did not.
	%
	%These divergent results raise the question if they are due to cross-linguistic variation, task differences, or artifacts of the analysis. Finally, much previous research calls into question the hypothesis that by-predicate differences in the effect of operator on projection pattern according to Kattunen's factive/semi-factive distinction (\citealt{beaver_have_2010,tonhauser_how_2018,sieker_projective_2022,degen_are_2022}).

    %Karttunen's distinction gives rise to three hypotheses about the projection associated with clause-embedding predicates, which we explore in this paper: \textbf{(i) by-operator variation:} entailment-cancelling operators differentially affect projection; \textbf{(ii) operator/predicate interaction:} the effect of operator differs between predicates; \textbf{(iii) factive/semi-factive distinction:} this difference can be addressed by characterizing lexical classes of (semi-)factive predicates.

    %To address the three hypotheses, we conducted a set of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family} for various clause-embedding predicates. We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' task; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English predicates, including factives and semi-factives, and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}). Our findings support (i) by-operator variation, and an operator/predicate interaction (ii), but not Karttunen's factive/semi-factive distinction (iii).

\section{Experiments}\label{s2}

	To assess the effect of entailment-cancelling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.
    %
    The predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal {\em perhaps} in Exps.~3, and in conditional antecedents in Exps.~4. Each set of experiments contained three experiments, that differed in an at-issueness measure used in a separate block. In this paper, we limit our attention to the projection ratings collected in these twelve experiments.\footnote{\label{footnote-supplement}The experiments, data and analysis scripts, as well as supplements referred to in this paper can be found in the following GitHub repository: \url{https://github.com/judith-tonhauser/CommitmentBankPlus}.}

	In all twelve experiments, projection was measured with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.} This diagnostic was also used in \citealt{sieker_projective_2022}. Under this diagnostic, participants are presented with utterances like those in \Next, and asked to rate whether the (named) speaker is certain of the CC.

@@

		\subsubsection{Participants}
			We recruited 250-300 participants for each of the 12 experiments. Participants for one experiment were recruited on Amazon's Mechanical Turk platform. These participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs. Participants for the remaining experiments were recruited on Prolific. These participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%. See Supplement \ref{app:d-participants} (in the repository linked to in footnote \ref{footnote-supplement}) for further information about the participants.

            \begin{figure}[h!]
				\centering
				\fbox{\includegraphics[width = .6\linewidth]{task-1n-proj.eps}}
				\caption{A sample trial in Exps.~2. In the corresponding trials in the other experiments, participants were presented with an utterance with a different entailment-cancelling operator.}
				\label{fig:trial}
			\end{figure}

@@

  %which were main clauses with appositive relative clauses, which were also presented as utterances made by a named speaker. Here, the main clause content was targeted by the projection question. In the experiments with polar questions, the controls were expected to receive low certainty ratings, whereas in the experiments with declaratives, they were expected to receive high ratings.

  %Each participant saw all 20 predicates, randomly paired with a unique embedded clause from a set of 20 contents, and a subject from a set of 20 proper names. %The predicates associated with projective content were embedded under (i) negation, (ii) polar questions, (iii) the epistemic possibility modal \textit{perhaps}, and (iv) in conditional antecedents in the four different sets of experiments.

        \subsubsection{Procedure}

			Participants were told to imagine that they are at a party and that, when walking into the kitchen, they overhear somebody say something to somebody else.
            %
            On each trial, participants read an utterance and responded to the `certain that'   question. They gave their response on a slider marked `no' (coded as 0) at one end and `yes' (coded as 1) on the other. A sample trial from Exps.~1 is given in Figure~\ref{fig:trial}.

            Following \citet{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things:
            %
            On the one hand, higher ratings could reflect greater speaker commitment towards the CC, and therefore greater projection. This assumes that speaker commitment is gradient.
            %
            Conversely, higher certainty ratings could reflect the probability that an interpreter takes the speaker to be committed to the CC.  On this interpretation, speaker commitment may be a binary, categorical property and projection variation is a result of the uncertainty about possible interpretations. %, which suggests the latter kind of interpretation). 
            In this paper, we remain agnostic about the underlying interpretation of projectivity as a gradient property, though our discussion of projection variability will be in line with the first interpretation (for discussion, see also \citealt{grove_factivity_2023}).

            %  we take high ratings of speaker certainty to reflect speaker greater speaker commitment towards the CC, that is, greater projection.

@@

        \subsubsection{Data exclusion}

        Data were excluded based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}. The data from 2,682 participants entered into the analysis.

            %We analyze the data from $2,682$ self-reported native speakers of American English, after excluding participants who failed one of the following two attention-checks: their mean rating on six control items is more than 2 sd different from the group mean, or they always roughly selected the same point on the scale for target stimuli (identified by low variance, and then manually inspecting their response patterns).

@@

        \mcdm{again needs to be consistent, sometimes Figure} Figure~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates. As shown, there is projection variability by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relatively more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.

			%When aggregating across predicates, we found an effect of \texttt{operator}. The distribution of projection ratings, as well as means and $95\%$ bootstrapped confidence intervals by operator are shown in Figure~\ref{fig:op-ratings}.

			\begin{figure}[ht]
				\centering
				\includegraphics[scale = .8]{operator-graph-1}
				\caption{Mean certainty ratings by operator. Error bars indicate $95\%$ bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the individual participants’ ratings.}
				\label{fig:op-ratings}
			\end{figure}

			%Mean projection ratings were higher for CCs in question-embeddings than under negation and modals, but lower than in conditional antecedents.

   These generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.\mcdm{we don't know what Model \#1 is}

			\begin{table}[ht]
					\centering
					\hspace{-1.3em}
					\begin{tabular}{llrrrr}
						Model & & Estimate & Std. Error & t-value\\
						\midrule
						\#1 & Intercept: \emph{question} & 0.51 & 0.01 & 44.78 & ***\\
						& operator: conditional & 0.05 & 0.01 & 5.30 & ***\\
						& operator: modal & -0.04 & 0.01 & -4.45 & ***\\
						& operator: negation & -0.03 & 0.01 & -4.67 & ***\\
						\bottomrule
					\end{tabular}

					\caption{Excerpt from the model output of a linear mixed effects model with fixed effects of operator; random effects: participant and item intercepts, fitted with \texttt{lme4, lmertest} in \texttt{R}.\label{t:op-model}}
				\end{table}

			Like \citet{sieker_projective_2022} for German, we observe a small main effect of entailment-cancelling operator on projection. However, in contrast to their result that projection is strongest from under negation, our results suggest that projection is strongest from antecedents of conditionals.\mcdm{does it make sense to compare different predicates?}
            We note, however, that the differences in mean certainty ratings between the four entailment-canceling operators are very small. This suggests that, when abstracting away from individual projective contents, projection from under the four entailment-canceling operators is very similar. In other words, when abstracting away from individual projective contents, the family of sentences really are a family.

@@

    Figure~\ref{fig:op-pred-ratings} shows mean certainty ratings by entailment-canceling operator for the five (semi-)factive predicates (indicated in pink), and the 15 non-factive predicates, where predicates are ordered by their overall mean certainty ratings.

    \begin{figure}[ht]
        \centering
		\includegraphics[width = \linewidth]{predicate-operator-graph-1}
		\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals. \textcolor{pred-fcv-color}{(Semi-)factive} and non-factive predicates are ordered by mean certainty rating pooled over the four operators.
        Embedding operator coded by letter and color:  \textcolor{op-n-color}{\texttt{N}: negation}, \textcolor{op-m-color}{\texttt{M}: modal \emph{perhaps}}, \textcolor{op-c-color}{\texttt{C}: conditional antecedents}, \textcolor{op-q-color}{\texttt{Q}: polar questions}.}
		\label{fig:op-pred-ratings}
	\end{figure}

    As shown, there is by-operator variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation. Further, the effect of operator differs between predicates. Examining the differences in certainty-ratings between the predicates for the four operators suggests that the observed variation does not align with widely assumed predicate classes. For instance, the five (semi-)factive predicates exhibit five different patterns:

    i) the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modal {\em perhaps}; ii) the CC of \emph{know} projects most from under questions, less from conditionals and negation, and least from under the modal {\em perhaps}; iii) the CC of \emph{see} projects most from under questions and conditionals, less from under negation, and least from under the modal \emph{perhaps}; iv) the CC of  \emph{discover} projects more from under conditionals and questions than negation and modal {\em perhaps}; and v) the CC of \emph{reveal} projects most from conditionals, less from questions, and least from negation and modal \emph{perhaps}. Thus, contrary to what \citealt{karttunen_observations_1971} proposed, there is no predicate whose CC projects uniformly from under all four entailment-canceling operators (as he suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators. We also observe by-operator projection variation for non-factives predicates. Some of this variation aligns with that observed for factive predicates: For instance, the CC of \emph{inform} exhibits the same pattern as the CC of \emph{know}, and the CC of \emph{acknowledge} exhibit the same pattern as that of the CC of \emph{discover}. Other non-factives predicates exhibit other patterns: For instance, the CCs of \emph{admit}, \emph{confess} and \emph{announce} project most from the antecedents of conditionals than the other three operators.

    These generalizations are confirmed by the statistical analysis. We fitted a Bayesian mixed-effects beta regression to the subset of the data containing the observations for each predicate, using the \texttt{brms} package (\citealt{burkner_brms_2017}) in R (\citealt{r_core_team_r_2014}). Each of the 20 models used low-information priors and estimated the effect of the \lq operator\rq\ (fixed effect) on scaled\footnote{The beta regression model assumes a beta-distribution defined over the open unit interval $(0,1)$, excluding boundaries. Following \citet{degen_are_2022}, we scaled the response variable, which is distributed over the closed interval $[0,1]$ (including boundaries), using the method in \citealt{smithson_better_2006}.} projection ratings, with random by-item intercepts (where an item is a complement clause), and by-item slopes for operators.

    For each predicate, the differences between operators were established by a pairwise comparison of posterior samples for the four operators using the \texttt{emmeans} package (\citealt{lenth_emmeans_2024}). This comparison yielded the $95\%$ high density interval (HDI) of estimated marginal mean differences for each pairwise combination of operators. We assume that two conditions differ when their HDI does not include $0$.

    The results are indicated in Figure~\ref{fig:op-pred-analysis}, which shows shows mean certainty ratings by entailment-canceling operator for the 20 predicates, here grouped by predicate. As in Figure~\ref{fig:op-pred-ratings}, (semi-)factive predicates are indicated in pink, and predicates are ordered by their overall mean certainty ratings. Non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet.

    \begin{figure}[t]
        \centering
		\includegraphics[width = \linewidth]{analysis.pdf}
		\caption{Mean certainty ratings by predicate (\textcolor{pred-fcv-color}{(semi-)factive}, non-factive) and operator (\textcolor{op-m-color}{modal},  \textcolor{op-n-color}{negation}, \textcolor{op-q-color}{polar questions}, \textcolor{op-c-color}{conditional antecedents}) with 95\% bootstrapped confidence intervals. Violin plots indicate kernel probability density of observations. Below each facet, a line spanning two operators indicates a non-zero difference according to the pairwise comparison of operators. The linetype indicates whether the difference is
        at least $1$ (solid line: —),
        below $1$ but at least $0.5$ (dashed line: – –),
        or below $0.5$ but above $0$ (dotted line: . . . )}
		\label{fig:op-pred-analysis}
	\end{figure}

    Because we found statistically robust non-zero differences by-operator differences for each of the predicates, and these differ beween predicates, the analysis also supports the more general claims that there is by-operator variation for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.

    As a result, our findings align with those of \citealt{smith_relationship_2014}, who also observed by-expression variation in the effect of operator. However, while they found that the CC of \emph{know} projects more from under negation than the antecedent of a conditional, we did not find an effect here. We hypothesize that this difference is due to the difference in projection diagnostic used.

	Our results differ from those of \citealt{sieker_projective_2022}: While their work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates. As \citealt{sieker_projective_2022} also used the `certain that' diagnostic for projection, this difference in results is not likely due to the diagnostic. Other factors that varied between our experiments are the language under investigated (German vs.\ English), the clause-embedding predicates investigated, and the CCs that the predicates were paired with. Future research will need to establish which of these factors are implicated in the observed differences.

@@

Factive predicates are defined as predicates that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker10,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} As presuppositions are assumed to project from under entail- ment-canceling operators, this definition leads one to expect that factive predicates are distinguished from non-factives ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives ones. This expectation was investigated in \citealt{degen_are_2022} based on the 20 clause-embedding predicates in \ref{ex:predicates} embedded in polar questions. Contrary to expectation,  \poscite{degen_are_2022} Exps.~1 found that the CCs of the five factive predicates varied in projection, that the CCs of the 15 non-factive predicates were projective compared to the non-projective main clause contents, and that the CCs of some non-factives predicates were as projective, or even more projective, than the CCs of some factive predicates.  In short, projection of the CC from under polar questions did not categorically distinguish factive from non-factives predicates. Further support for this result came from the analysis of projection ratings in three additional datasets, namely the CommitmentBank (\citealt{demarneffe-etal-sub23}), the VerbVeridicality dataset (\citealt{ross-pavlick2019}), and the MegaVeridicality  dataset (\citealt{white_role_2018}).

@@

>>>>

%% other packages I added (LH)
% misc formatting
	\usepackage{booktabs}
    \usepackage[export]{adjustbox}
    \usepackage{rotating}
    \usepackage{longtable}

@@

    Some research, however, suggests that entailment-cancelling operators may affect projection differentially. For instance, \citet{karttunen_observations_1971} proposed distinguishing English factive predicates (e.g., \textit{regret}) from semi-factives (e.g., \textit{discover}). Based on \Next, he argued that the content of the complement (CC) of true factives consistently projects across the operators in \Last, but that of semi-factives does not always project from under polar questions, modals, or conditionals.%

@@

\section{Experiments}\label{s2}
	To assess the effect of entailment-cancelling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.
    %
    The predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal {\em perhaps} in Exps.~3, and in conditional antecedents in Exps.~4. Each set of experiments contained three experiments, that differed in an at-issueness measure used in a separate block. In this paper, we limit our attention to the projection ratings collected in these twelve experiments.\footnote{\label{footnote-supplement}The experiments, data and analysis scripts, as well as supplements referred to in this paper can be found in the following GitHub repository: \url{https://github.com/judith-tonhauser/CommitmentBankPlus}.}

	In all twelve experiments, projection was measured with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019,sieker_projective_2022}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.}
    %This diagnostic was also used in \citealt{sieker_projective_2022}.
    Under this diagnostic, participants are presented with utterances like those in \Next, and asked to rate whether the (named) speaker is certain of the CC.

@@

		%\subsubsection{Participants}
        \paragraph{Participants.}
			We recruited 250-300 participants for each of the 12 experiments. Participants for one experiment were recruited on Amazon's Mechanical Turk platform. These participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs. Participants for the remaining experiments were recruited on Prolific. These participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%. See Supplement \ref{app:d-participants} (in the repository linked to in footnote \ref{footnote-supplement}) for further information about the participants.

            \begin{figure}[h!]
				\centering
				\fbox{\includegraphics[width = .6\linewidth]{task-1n-proj.eps}}
				\caption{A sample trial in Exps.~1. In the corresponding trials in the other experiments, participants were presented with an utterance with a different entailment-cancelling operator.}
				\label{fig:trial}
			\end{figure}

@@

        \subsubsection{Procedure}
            Participants were asked to imagine that they are at a party and that, when walking into the kitchen, they overhear somebody say something to somebody else.
            %
            On each trial, they read an utterance and gave a response to the `certain that' question on a slider marked `no' (coded as 0) at one end and `yes' (coded as 1) on the other. A sample trial from Exps.~1 is shown in Figure~\ref{fig:trial}.

            Following \citet{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things: On the one hand, higher ratings could reflect greater speaker commitment towards the CC, and therefore greater projection. This assumes that speaker commitment is interpreted in a gradient way.
            % gradient.
            Conversely, higher certainty ratings could reflect the probability that an interpreter takes the speaker to be committed to the CC.  On this interpretation, speaker commitment may be a binary, categorical property and projection variation is a result of the uncertainty about possible interpretations. %, which suggests the latter kind of interpretation). 
            In this paper, we remain agnostic about the underlying interpretation of projectivity as a gradient property, though our discussion of projection variability will be in line with the first interpretation (for discussion, see also \citealt{grove_factivity_2023}).

@@

        \subsubsection{Data exclusion}
            Data were excluded based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}. The data from 2,682 participants entered into the analysis.

@@

        Figure~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates. As shown, there is projection variability by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relatively more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.

            \begin{figure}[ht]
				\centering
				\includegraphics[scale = .8]{certainty-operator.pdf}
				\caption{Mean certainty ratings by operator. Error bars indicate $95\%$ bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the ratings.
                % of the individual participants’ ratings.
                }
				\label{fig:op-ratings}
			\end{figure}

   These observations were confirmed by a posthoc pairwise comparison of the estimated means for each entailment-canceling operator using the `emmeans' package (\citealt{emmeans}) in R (\citealt{r}). The input to the pairwise comparison was a  Bayesian mixed-effects beta regression model with weakly informative priors that were fit using the \emph{brms} package (\citealt{buerkner2017}). The model predicted certainty ratings\footnote{To model the certainty ratings using a beta regression, the ratings were first transformed from the interval [0,1] to the interval (0,1) using the method proposed in \citealt{smithson-verkuilen2006}.} from a fixed effect of entailment-canceling operator (with treatment coding and `modal' as the reference level and included a random by-predicate intercept. The output of the pairwise comparison were 95\% highest density intervals (HDIs) of estimated marginal mean differences between each of the operators. We assume that two operators differ in certainty ratings if their HDI does not include 0.\footnote{A more detailed report of the model output is provided in Supplement \ref{app:analysis-1}}

Table \ref{t:pairwise} provides the output of the pairwise comparisons. As shown,

, the pairwise comparison suggests differences between the naturalness ratings of the CCs of the five factive predicates: The ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}.\footnote{While the naturalness ratings collected from the participants range from 0 to 1, the greater than $|$1$|$  estimated marginal mean differences are the result of the beta regression using a logit link function for the mean parameter.} There is also variation between the factive predicates and the presuppositional controls: While all five factive predicates are more natural in explicit ignorance contexts than {\em continue, too} and {\em also}, {\em know} is additionally more natural than {\em again} and {\em stop}, and {\em see, discover} and {\em reveal} are additionally more natural than the {\em it-}cleft. Finally, the naturalness ratings in explicit ignorance contexts for {\em see} and {\em discover} do not differ from those of the nonfactives {\em confess, inform}, and {\em acknowledge}, and those of {\em reveal} do not differ from those of the nonfactives {\em admit, hear}, and {\em prove}. These results suggest that naturalness ratings in explicit ignorance contexts do not provide empirical support for categorical projection or for a categorical distinction between factive and nonfactive predicates.

\begin{table}[!h]
\centering
\begin{tabular}{lrrr}
 contrast & estimate & lower 95\% CI & upper 95\% CI \\\hline 
 conditional - negation & 0.21 & 0.18 & 0.24 \\ 
  conditional - question & 0.13 & 0.10 & 0.16 \\ 
  modal - conditional & -0.26 & -0.29 & -0.23 \\ 
  modal - negation & -0.05 & -0.08 & -0.02 \\ 
  modal - question & -0.14 & -0.17 & -0.11 \\ 
  negation - question & -0.09 & -0.12 & -0.06 \\ 
\end{tabular}
\caption{Pairwise differences between entailment-canceling operators.}\label{t:pairwise}
\end{table}

		Like \citet{sieker_projective_2022} for German, we observe a small main effect of entailment-cancelling operator on projection. However, in contrast to their result that projection is strongest from under negation, our results suggest that projection is strongest from antecedents of conditionals.
        % \mcdm{does it make sense to compare different predicates?}
        We note, however, that the differences in mean certainty ratings between the four entailment-canceling operators are very small. This suggests that, when abstracting away from individual projective contents, projection from under the four entailment-canceling operators is very similar. In other words, when abstracting away from individual projective contents, the family of sentences really are a family.

@@

        Figure~\ref{fig:op-pred-analysis} shows mean certainty ratings by entailment-canceling operator for the five (semi-)factive predicates, and the 15 non-factive predicates, where predicates are ordered by their overall mean certainty ratings.
        %
        As shown, there is by-operator variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation. Further, the effect of operator differs between predicates. Examining the observed variation suggests that it does not align with widely assumed predicate classes.

         \begin{figure}[ht!]
            \centering
    		\includegraphics[width = .9\linewidth]{analysis.pdf}
    		\caption{Mean certainty ratings by predicate (\textcolor{pred-fcv-color}{(semi-)factive}, non-factive) and operator (\textcolor{op-m-color}{modal},  \textcolor{op-n-color}{negation}, \textcolor{op-q-color}{polar questions}, \textcolor{op-c-color}{conditional antecedents}) with 95\% bootstrapped confidence intervals. Violin plots indicate kernel probability density of the individual participants' ratings. Facets for predicates are ordered by the predicate mean certainty rating (aggregating across operators). Below each facet, a line spanning two operators indicates a non-zero difference according to the pairwise comparison of operators. The linetype indicates whether the difference is
            at least $1$ (solid line: —),
            below $1$ but at least $0.5$ (dashed line: – –),
            or below $0.5$ but above $0$ (dotted line: . . . )}
    		\label{fig:op-pred-analysis}
    	\end{figure}

        For instance, the five (semi-)factive predicates exhibit four different patterns:
        %
        First, the CC of \emph{be annoyed} projects
        most from under questions, less from under negation, followed by conditionals, and least from under the modal \emph{perhaps} (\textcolor{op-q-color}{Q} $>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-c-color}{C} $>$ \textcolor{op-m-color}{M}).
        % more from under negation and questions than conditionals and the modal {\em perhaps};
        Second, the CC of \emph{know} projects most from under questions, less from conditionals and negation, and least from under {\em perhaps} (\textcolor{op-q-color}{Q} $> \{$\textcolor{op-n-color}{N}$,$ \textcolor{op-c-color}{C}$\} >$ \textcolor{op-m-color}{M}).
        %
        The CCs of \emph{discover} and \emph{see} exhibit a third pattern: They project most from under questions and conditionals, less from under negation, and least from under \emph{perhaps} ($\{$\textcolor{op-q-color}{Q}$,$ \textcolor{op-c-color}{C}$\}>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-m-color}{M}).
        %
        Finally, the CC of \emph{reveal} projects most from conditionals, less from questions, and least from under negation and \emph{perhaps} (\textcolor{op-c-color}{C} $>$ \textcolor{op-q-color}{Q}  $> \{$\textcolor{op-n-color}{N}$,$ \textcolor{op-m-color}{M}$\}$).
        %
        Thus, contrary to what \citealt{karttunen_observations_1971} proposed, there is no predicate whose CC projects uniformly from under all four entailment-canceling operators (as he suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators.

        We also observe by-operator projection variation for non-factives predicates. Some of this variation aligns with that observed for factive predicates: For instance, the CC of \emph{inform} exhibits the same pattern as the CC of \emph{know}, and the CC of \emph{acknowledge} exhibit the same pattern as that of the CC of \emph{discover}. Other non-factives predicates exhibit other patterns: For instance, the CCs of \emph{admit}, \emph{confess} and \emph{announce} project most from the antecedents of conditionals than the other three operators.

        These generalizations are confirmed by the statistical analysis. We fitted a Bayesian mixed-effects beta regression to the subset of the data containing the observations for each predicate (using \texttt{brms} in R). Each of the 20 models used low-information priors and estimated the effect of the \lq operator\rq\ (fixed effect) on (scaled) projection ratings, with random by-item intercepts and operator slopes. For each predicate, the differences between operators were established by a pairwise comparison of posterior estimates of marginal means for the four operators (using \texttt{emmeans}). We take those differences between posterior samples to be statistically robust where the $95\%$ (HDI) does not include $0$.\footnote{Further details about the model are provided in Supplement \ref{app:analysis-2}.} The results are indicated in Figure~\ref{fig:op-pred-analysis}, where non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet.

        Because we found statistically robust non-zero differences by-operator differences for each of the predicates, and these differ between predicates, the analysis also supports the more general claims that there is by-operator variation for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.
        %
        Therefore, our findings align with those of \citealt{smith_relationship_2014}, who also observed by-expression variation in the effect of operator. However, while they found that the CC of \emph{know} projects more from under negation than the antecedent of a conditional, we did not find an effect here. We hypothesize that this difference is due to the difference in projection diagnostic used.

        Our results, however, differ from those of \citealt{sieker_projective_2022}: While their work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates. As \citealt{sieker_projective_2022} also used the `certain that' diagnostic for projection, this difference in results is not likely due to the diagnostic. Other factors that varied between our experiments are the language under investigated (German vs.\ English), the clause-embedding predicates investigated, and the CCs that the predicates were paired with. Future research will need to establish which of these factors are implicated in the observed differences.

@@

        Factive predicates are defined as predicates that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} As presuppositions are assumed to project from under entail- ment-canceling operators, this definition leads one to expect that factive predicates are distinguished from non-factives ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives ones. This expectation was investigated in \citealt{degen_are_2022} based on the 20 clause-embedding predicates in \ref{ex:predicates} embedded in polar questions. Contrary to expectation,  \poscite{degen_are_2022} Exps.~1 found that the CCs of the five factive predicates varied in projection, that the CCs of the 15 non-factive predicates were projective compared to the non-projective main clause contents, and that the CCs of some non-factives predicates were as projective, or even more projective, than the CCs of some factive predicates.  In short, projection of the CC from under polar questions did not categorically distinguish factive from non-factives predicates. Further support for this result came from the analysis of projection ratings in three additional datasets, namely the CommitmentBank (\citealt{demarneffe-etal-sub23}), the VerbVeridicality dataset (\citealt{ross-pavlick2019}), and the MegaVeridicality  dataset (\citealt{white_role_2018}).

@@

\section{Analysis 1 --- Estimating operator effects on projection} \label{app:analysis-1}

    Bayesian mixed effects beta regression for the full dataset.\footnote{The analysis script can be found at \url{https://github.com/judith-tonhauser/CommitmentBankPlus/blob/main/results/main/meta-analyses/1_projection/rscripts/2_analysis-operators.R}.}\\

    \noindent\textbf{Repsonse variable:} projection ratings scaled for beta-regression to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$,
    % using method used in \cite{degen_are_2022}, from \cite{smithson_better_2006}, for proportional data.
    $$y' = \frac{y * (n - 1) + 0.5}{n}$$

    \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)\\

    \noindent\textbf{Random effects:}

\section{Analysis 2 --- Estimating operator effects for each predicate} \label{app:analysis-2}

    Bayesian mixed effects beta regression for the each subset of the dataset by predicate.\\

    \noindent\textbf{Repsonse variable:} projection ratings scaled for beta-regression.\\

    \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)\\

    \noindent\textbf{Random effects:} Item random intercepts and slopes by operator; No participant random effects, because each participant saw each predicate only once.\\

    \noindent The models were fit using \texttt{brms}
    % (\citealt{burkner_brms_2017}) in R (\citealt{r_core_team_r_2014}), 
    using low-information priors (\texttt{brms} defaults). The estimation used $4$ chains, $4,000$ iterations per chain, $700$ warm-up, total $13,200$ iterations saved for each model.

    \subsection{Model output}

    For each predicate, differences between levels of operator were established by considering the pairwise difference between estimated marginal means for each level of operator, using the \texttt{emmeans} package (\citealt{lenth_emmeans_2024}) in R (R Core Team 2016). The below table gives posterior means, as well as upper and lower values for 95\% highest density intervals (HDIs) for the derived variables representing the pairwise differences.
